---
title: "HW1 GR5206"
author: Tianchen Wang / tw2665
date: Sep 14, 2018
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### Part 1
i)
```{r Import Data}
titanic <- read.table('Titanic.txt', as.is = TRUE, header = TRUE, sep = '')
```

ii)
```{r}
dim(titanic)
```
891 rows and 12 columns

iii)
```{r Survived.Word}
died <- which(titanic$Survived == 0)
Survived.Word <- rep('survived', dim(titanic)[1])
Survived.Word[died] <- 'died'
titanic$Survived.Word <- Survived.Word
```
### Part 2
i)
```{r Exploring Data}
dataForMean <- cbind(titanic$Survived, titanic$Age, titanic$Fare)
meanForEachCol <- apply(dataForMean, 2, mean)
# check if there are NA in Age column
meanForEachCol
sum(is.na(titanic$Age)) > 0
```
The mean of age column is NA, because there are NA in this column and such data is not a numeric data, which means the function **mean** does not know how to deal with it without setting **na.rm == TRUE**.

ii)
```{r}
totalNum <- dim(titanic)[1]
femaleSurvived <- titanic[titanic$Sex == "female", ]
femaleNum <- dim(femaleSurvived)[1]
femaleSurvived <- femaleSurvived[femaleSurvived$Survived == 1,]
femaleSurvivedNum <- dim(femaleSurvived)[1]
round(femaleSurvivedNum / femaleNum, digits = 2)
round(femaleSurvivedNum / totalNum, digits = 2)

```
Since I am not sure what is meaning of "This is not a conditional probability" in the question. I just compute the proportion of both survived female out of the whole titanic victim list and the proportion of survived female out of the female passengers.
The first one is 0.26. The second one is 0.74, which means most female were survived. I have to say this really showed the good in human nature of the passangers of Titanic. 

iii)
```{r}
survived <- titanic[titanic$Survived == 1, ]
survivedNum <- dim(survived)[1]
round(femaleSurvivedNum / survivedNum, digits = 2)

```

iv)
```{r}
classes <- sort(unique(titanic$Pclass))
Pclass.Survival <- vector("numeric", length = 3)
names(Pclass.Survival) <- classes

for (i in 1:3) {
  temData <- titanic[titanic$Pclass == i, ]
  proportion <- mean(temData$Survived)
  Pclass.Survival[i] <- round(proportion, digits = 2)
}
remove(temData)
Pclass.Survival
```

v)
```{r}
Pclass.Survival2 <- tapply(titanic$Survived, titanic$Pclass, 
                           function(x) round(mean(x), digits = 2))
Pclass.Survival2
```
Basically, the **tapply** will apply function on data in the first input separated by the different level of second input. 

vi)
From the survived rate computed above, it is easy to find the higher the class rank is, the higher the survival rate.

