---
title: "Lab 1"
author: "Tianchen Wang / tw2665"
date: "September 9, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r, echo = FALSE}
set.seed(1) # Please don't remove this code!
```

# Instructions 
Before you leave lab today make sure that you upload an RMarkdown file to the canvas page (this should have a .Rmd extension) as well as the html output after you have knitted the file (this will have a .html extension).  Note that since you have already knitted this file, you should see both a **Lab1_UNI.html** and a **Lab1_UNI.Rmd** file in your GR5206 folder.  Click on the **Files** tab to the right to see this.  The files you upload to the Canvas page should be updated with commands you provide to answer each of the questions below.  You can edit this file directly to produce your final solutions. 

# Background: The Normal Distribution
Recall from your probability class that a random variable $X$ is normally-distributed with mean $\mu$ and variance $\sigma^2$ (denoted $X \sim N(\mu, \sigma^2)$) if it has a probability density function, or *pdf*, equal to

\[f(x) = \frac{1}{\sqrt{2\pi \sigma^2}} e^{-\frac{(x - \mu)^2}{2\sigma^2}}.\]

In *R* we can simulate $N(\mu, \sigma^2)$ random variables using the \texttt{rnorm()} function.  For example,

```{r}
rnorm(n = 5, mean = 10, sd = 3)
```

outputs 5 normally-distributed random variables with mean equal to 10 and standard deviation (this is $\sigma$) equal to 3.  If the second and third arguments are ommited the default rates are **mean = 0** and **sd = 1**, which is referred to as the "standard normal distribution".

# Tasks

## Sample means as sample size increases

1) Generate 100 random draws from the standard normal distribution and save them in a vector named **normal100**.  Calculate the mean and standard deviation of **normal100**.  In words explain why these values aren't exactly equal to 0 and 1.

```{r}
# You'll want to type your response here.  Your response should look like:
normal100 <- rnorm(n = 100, mean = 0, sd = 1)
mean(normal100)
sd(normal100)
# Of course, your answer should not be commented out.
```
The reason why mean and standard deviation of **normal100** are not exactly 0 and 1 is that we randomly draw out 100 numbers by using the standard normal distribution. These numbers are independent, which means that every number in **normal100** should fall in the real number space and the probability of which number it will be is determined by standard normal distribution but what exactly the value is unknown. So the mean and sd of **normal100** may be 0 and 1 or may be not. But we could expect that mean and sd will be close to 0 and 1.

2) The function **hist()** is a base *R* graphing function that plots a histogram of its input.  Use **hist()** with your vector of standard normal random variables from question (1) to produce a histogram of the standard normal distribution.  Remember that typing **?hist** in your console will provide help documents for the **hist()** function.  If coded properly, these plots will be automatically embedded in your output file.
```{r}
hist(normal100)
```


3) Repeat question (1) except change the number of draws to 10, 1000, 10,000, and 100,000 storing the results in vectors called **normal10**, **normal1000**, **normal10000**, **normal100000**. 
```{r}
normal10 <- rnorm(10)
normal1000 <- rnorm(1000)
normal10000 <- rnorm(10000)
normal100000 <- rnorm(100000)

```


4) We want to compare the means of our four random draws.  Create a vector called **sample_means** that has as its first element the mean of **normal10**, its second element the mean of **normal100**, its third element the mean of **normal1000**, its fourth element the mean of **normal10000**, and its fifth element the mean of **normal100000**.  After you have created the **sample_means** vector, print the contents of the vector and use the **length()** function to find the length of this vector.  (it should be five).  There are, of course, multiple ways to create this vector.  Finally, explain in words the pattern we are seeing with the means in the **sample_means** vector.
```{r}
sample_means <- c(mean(normal10), mean(normal100), mean(normal1000),
                 mean(normal10000),mean(normal100000))
length(sample_means)
sample_means
```
It is easy to find the number in **sample_means** is getting closer to 0 from the mean of first vector(**normal10**) to last vector(**normal100000**). The reason behind this phenomena is the Law of Large Numbers, which indicates that the mean of r.v. is getting closer to its expection when number of sample is getting larger and larger.

## Sample distribution of the sample mean

5)  Let's push this a little farther.  Generate 1 million random draws from a normal distribution with $\mu = 3$ and $\sigma^2 = 4$ and save them in a vector named **normal1mil**.  Calculate the mean and standard deviation of **normal1mil**.
```{r}
normal1mil <- rnorm(n = 1000000, mean = 3, sd = sqrt(4))
mean(normal1mil)
sd(normal1mil)
```

6) Find the mean of all the entries in **normal1mil** that are greater than 3.  You may want to generate a new vector first which identifies the elements that fit the criteria.
```{r}
gter3_pos <- which(normal1mil > 3)
gter3 <- normal1mil[gter3_pos]
mean(gter3)
```


7) Create a matrix **normal1mil_mat** from the vector **normal1mil** that has 10,000 columns (and therefore should have 100 rows).
```{r}
normal1mil_mat <- matrix(normal1mil, nrow = 100, ncol = 10000)

```


8) Calculate the mean of the $1234^{th}$ column.
```{r}
mean(normal1mil_mat[,1234])
# apply(normal1mil_mat[,c(1,2,3,4)], MARGIN = 2, FUN = mean)
```


9) Use the **colSums()** functions to calculate the *means* of each column of **normal1mil_mat**.  Remember, **?colSums** will give you help documents about this function.  Save the vector of column means with an appropriate name as it will be used in the next task.
```{r}
# since there are 100 rows in the normal1mi_mat
col_means = colSums(normal1mil_mat, dims = 1)/100
```


10) Finally, produce a histogram of the column means you calculated in task (9).  What is the distribution that this histogram approximates (i.e.\ what is the distribution of the sample mean in this case)?
```{r}
hist(col_means)
```

It looks like a normal distribution with mean 3.


